---
title: "ScleriaBiogeography2"
author: "amesclir"
date: "12/03/2019"
output: html_document
editor_options: 
chunk_output_type: console
---

Installing last version of BioGeoBEARS from Github

```{r, echo=F}
# Only use these codes if you have not installed BioGeoBEARS
#install.packages("rexpokit")
#install.packages("cladoRcpp")

#library(phylobase)
#library(pillar)
#library(RNeXML)

#library(devtools)
#devtools::install_github(repo="liamrevell/phytools") #Installing last version of phytools from Github

#devtools::install_github(repo="nmatzke/BioGeoBEARS") #Installing last version of BioGeoBEARS from Github
```


First, load the packages.

```{r, echo=F}
library(ape)
library(phytools)
library(BioGeoBEARS)
library(parallel)
library(minqa)
library(optimx)

#NOTE: The BioGeoBEARS from GitHub must be installed (above), version 1.1.1
```

First, load the tree and distribution.

```{r, echo=F}

mytree1 <- read.nexus("tree.tree")
mytree2 <- drop.tip(mytree1, c("Becquerelia_cymosa_SC281", "Bisboeckelera_microcephala_SC282", "Calyptrocarya_glomerulata_SC106", "Diplacrum_caricinum_SC279"))

write.tree(mytree2, "tree.tree")

tr = read.tree("tree.tree")
tr
plot(tr, cex = 0.4)
title("Example Scleria phylogeny")
axisPhylo()

trfn = np("tree.tree")
moref(trfn)

geogfn = np("Scleria_BGB_occurrence_world.txt")
#extdata_dir = np(system.file("extdata", package="BioGeoBEARS"))
#extdata_dir
#list.files(extdata_dir)
#geogfn = np(paste(addslash(extdata_dir), "Psychotria_geog.data", sep=""))

geogfn
moref(geogfn)

# Look at your geographic range data:
tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
tipranges

max_range_size = 7

numstates_from_numareas(numareas=7, maxareas=7, include_null_range=TRUE)
#numstates_from_numareas(numareas=8, maxareas=8, include_null_range=FALSE)


```

3. Ahora vamos a usar el modelo DEC

```{r, echo=F}

#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Scleria_DEC_M0_unconstrained_v1.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDEC = res
    } else {
    # Loads to "res"
    load(resfn)
    resDEC = res
    }



```

4. Run DEC + J

```{r, echo=F}



#######################################################
# Run DEC+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDEC$outputs@params_table["d","est"]
estart = resDEC$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Add j as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
resfn = "Scleria_DEC+J_M0_unconstrained_v1.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDECj = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECj = res
    }

```


5. Plotting the results

```{r, echo=F}

#######################################################
# PDF plots
#######################################################
pdffn = "Scleria_DEC_vs_DEC+J_M0_unconstrained_v1.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC on Scleria M0_unconstrained"

# Setup
results_object = resDEC
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DECJ
#######################################################
analysis_titletxt ="BioGeoBEARS DEC+J on Scleria M0_unconstrained"

# Setup
results_object = resDECj
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()  # Turn off PDF



```

6. Run DIVA like

```{r, echo=F}
#######################################################
# Run DIVALIKE
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE model
# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# No jump dispersal/founder-event speciation
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.01
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.01

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

runslow = TRUE
resfn = "Scleria_DIVALIKE_M0_unconstrained_v1.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDIVALIKE = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKE = res
    }

```

7. Run DIVALIKE + j

```{r, echo=F}
#######################################################
# Run DIVALIKE+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDIVALIKE$outputs@params_table["d","est"]
estart = resDIVALIKE$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add jump dispersal/founder-event speciation
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Under DIVALIKE+J, the max of "j" should be 2, not 3 (as is default in DEC+J)
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","min"] = 0.00001
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","max"] = 1.99999

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

resfn = "Scleria_DIVALIKE+J_M0_unconstrained_v1.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDIVALIKEj = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEj = res
    }


```

8. Plotting additional results

```{r, echo=F}

load("Scleria_DIVALIKE_M0_unconstrained_v1.Rdata")
resDIVALIKE = res
load("Scleria_DIVALIKE+J_M0_unconstrained_v1.Rdata")
resDIVALIKEj = res

pdffn = "Scleria_DIVALIKE_vs_DIVALIKE+J_M0_unconstrained_v1.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DIVALIKE
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE on Scleria M0_unconstrained"

# Setup
results_object = resDIVALIKE
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DIVALIKE+J
#######################################################
analysis_titletxt ="Scleria DIVALIKE+J on Scleria M0_unconstrained"

# Setup
results_object = resDIVALIKEj
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()



```



9. Comparing all results

```{r, echo=F}
#########################################################################
#########################################################################
#########################################################################
#########################################################################
# 
# CALCULATE SUMMARY STATISTICS TO COMPARE
# DEC, DEC+J, DIVALIKE, DIVALIKE+J
# 
#########################################################################
#########################################################################
#########################################################################
#########################################################################

#########################################################################
#########################################################################
# REQUIRED READING:
#
# Practical advice / notes / basic principles on statistical model 
#    comparison in general, and in BioGeoBEARS:
# http://phylo.wikidot.com/advice-on-statistical-model-comparison-in-biogeobears
#########################################################################
#########################################################################

# Set up empty tables to hold the statistical results
restable = NULL
teststable = NULL

#######################################################
# Statistics -- DEC vs. DEC+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDEC)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDECj)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DEC, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDEC, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DEC+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECj, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

# The null hypothesis for a Likelihood Ratio Test (LRT) is that two models
# confer the same likelihood on the data. See: Brian O'Meara's webpage:
# http://www.brianomeara.info/tutorials/aic
# ...for an intro to LRT, AIC, and AICc

rbind(res2, res1)
tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#######################################################
# Statistics -- DIVALIKE vs. DIVALIKE+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKE)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEj)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DIVALIKE, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKE, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DIVALIKE+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEj, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

rbind(res2, res1)
conditional_format_table(stats)

tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)


#########################################################################
# RESULTS: DEC, DEC+J, DIVALIKE, DIVALIKE+J
#########################################################################
teststable$alt = c("DEC+J", "DIVALIKE+J")
teststable$null = c("DEC", "DIVALIKE")
row.names(restable) = c("DEC", "DEC+J", "DIVALIKE", "DIVALIKE+J")

# Look at the results!!
restable
teststable

#######################################################
# Save the results tables for later -- check for e.g.
# convergence issues
#######################################################

# Loads to "restable"
save(restable, file="restable_v1.Rdata")
load(file="restable_v1.Rdata")

# Loads to "teststable"
save(teststable, file="teststable_v1.Rdata")
load(file="teststable_v1.Rdata")

# Also save to text files
write.table(restable, file="restable.txt", quote=FALSE, sep="\t")
write.table(unlist_df(teststable), file="teststable.txt", quote=FALSE, sep="\t")

#######################################################
# Model weights of all four models
#######################################################
restable2 = restable

# With AICs:
AICtable = calc_AIC_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams)
restable = cbind(restable, AICtable)
restable_AIC_rellike = AkaikeWeights_on_summary_table(restable=restable, colname_to_use="AIC")
restable_AIC_rellike

# With AICcs -- factors in sample size
samplesize = length(tr$tip.label)
AICtable = calc_AICc_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams, samplesize=samplesize)
restable2 = cbind(restable2, AICtable)
restable_AICc_rellike = AkaikeWeights_on_summary_table(restable=restable2, colname_to_use="AIC")
restable_AICc_rellike
free_params = row.names(resDECj$output@params_table[resDECj$output@params_table$type=="free",])
names(restable_AICc_rellike) = c("LnL", "numparams", free_params, "AICc", "AICc_wt")

# Also save to text files
write.table(restable_AIC_rellike, file="restable_AIC_rellike.txt", quote=FALSE, sep="\t")
write.table(restable_AICc_rellike, file="restable_AICc_rellike.txt", quote=FALSE, sep="\t")

# Save with nice conditional formatting
write.table(conditional_format_table(restable_AIC_rellike), file="restable_AIC_rellike_formatted.txt", quote=FALSE, sep="\t")
write.table(conditional_format_table(restable_AICc_rellike), file="restable_AICc_rellike_formatted.txt", quote=FALSE, sep="\t")


```


End of the first approach: DEC, DEC+J, DIVALIKE and DIVALIKE+J without dispersal multipliers.




10. Now we are going to use the model DEC with dispersal multipliers allowing the dispersal matrix to be modified by extra parameter -w

```{r, echo=F}

#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = areasadjacency
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Scleria_DEC_M0_constrained_w.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDECc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECc = res
    }



```



11. Run DEC + J con tasas de dispersion

```{r, echo=F}



#######################################################
# Run DEC+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDEC$outputs@params_table["d","est"]
estart = resDEC$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Add j as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
resfn = "Scleria_DEC+J_M0_constrained_w.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDECjc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECjc = res
    }
```


12. Plotting the results with dispesal rates

```{r, echo=F}

#######################################################
# PDF plots
#######################################################
pdffn = "Scleria_DEC_vs_DEC+J_M0_constrained_w.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC on Scleria M0_constrained"

# Setup
results_object = resDECc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DECJ
#######################################################
analysis_titletxt ="BioGeoBEARS DEC+J on Scleria M0_constrained"

# Setup
results_object = resDECjc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()  # Turn off PDF

 # Plot it

```

13. Run DIVA like with dispersal rates

```{r, echo=F}
#######################################################
# Run DIVALIKE
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE model
# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

# No jump dispersal/founder-event speciation
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.01
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.01

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

runslow = TRUE
resfn = "Scleria_DIVALIKE_M0_constrained_w.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDIVALIKEc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEc = res
    }

```

14. Run DIVALIKE + j with dispersal rates

```{r, echo=F}
#######################################################
# Run DIVALIKE+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDIVALIKE$outputs@params_table["d","est"]
estart = resDIVALIKE$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add jump dispersal/founder-event speciation
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Under DIVALIKE+J, the max of "j" should be 2, not 3 (as is default in DEC+J)
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","min"] = 0.00001
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","max"] = 1.99999

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

resfn = "Scleria_DIVALIKE+J_M0_constrained_w.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDIVALIKEjc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEjc = res
    }


```

15. Plotting additional results with dispersal rates

```{r, echo=F}

pdffn = "Scleria_DIVALIKE_vs_DIVALIKE+J_M0_constrained_w.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DIVALIKE
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE on Scleria M0_constrained"

# Setup
results_object = resDIVALIKEc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DIVALIKE+J
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE+J on Scleria M0_constrained"

# Setup
results_object = resDIVALIKEjc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()



```



16. Comparing all results with dispersal rates

```{r, echo=F}
#########################################################################
#########################################################################
#########################################################################
#########################################################################
# 
# CALCULATE SUMMARY STATISTICS TO COMPARE
# DEC, DEC+J, DIVALIKE, DIVALIKE+J
# 
#########################################################################
#########################################################################
#########################################################################
#########################################################################

#########################################################################
#########################################################################
# REQUIRED READING:
#
# Practical advice / notes / basic principles on statistical model 
#    comparison in general, and in BioGeoBEARS:
# http://phylo.wikidot.com/advice-on-statistical-model-comparison-in-biogeobears
#########################################################################
#########################################################################

# Set up empty tables to hold the statistical results
restable = NULL
teststable = NULL

#######################################################
# Statistics -- DEC vs. DEC+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDECc)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDECjc)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DEC, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DEC+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECjc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

# The null hypothesis for a Likelihood Ratio Test (LRT) is that two models
# confer the same likelihood on the data. See: Brian O'Meara's webpage:
# http://www.brianomeara.info/tutorials/aic
# ...for an intro to LRT, AIC, and AICc

rbind(res2, res1)
tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#######################################################
# Statistics -- DIVALIKE vs. DIVALIKE+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEc)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEjc)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DIVALIKE, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DIVALIKE+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEjc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

rbind(res2, res1)
conditional_format_table(stats)

tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#########################################################################
# RESULTS: DEC, DEC+J, DIVALIKE, DIVALIKE+J
#########################################################################
teststable$alt = c("DEC+Jc", "DIVALIKE+Jc")
teststable$null = c("DECc", "DIVALIKEc")
row.names(restable) = c("DECc", "DEC+Jc", "DIVALIKEc", "DIVALIKE+Jc")

# Look at the results!!
restable
teststable

#######################################################
# Save the results tables for later -- check for e.g.
# convergence issues
#######################################################

# Loads to "restable"
save(restable, file="restable_wc.Rdata")
load(file="restable_wc.Rdata")

# Loads to "teststable"
save(teststable, file="teststable_wc.Rdata")
load(file="teststable_wc.Rdata")

# Also save to text files
write.table(restable, file="restablewc.txt", quote=FALSE, sep="\t")
write.table(unlist_df(teststable), file="teststablewc.txt", quote=FALSE, sep="\t")

#######################################################
# Model weights of all six models
#######################################################
restable2 = restable

# With AICs:
AICtable = calc_AIC_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams)
restable = cbind(restable, AICtable)
restable_AIC_rellike = AkaikeWeights_on_summary_table(restable=restable, colname_to_use="AIC")
restable_AIC_rellike

# With AICcs -- factors in sample size
samplesize = length(tr$tip.label)
AICtable = calc_AICc_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams, samplesize=samplesize)
restable2 = cbind(restable2, AICtable)
restable_AICc_rellike = AkaikeWeights_on_summary_table(restable=restable2, colname_to_use="AIC")
restable_AICc_rellike
free_params = row.names(resDECj$output@params_table[resDECj$output@params_table$type=="free",])
names(restable_AICc_rellike) = c("LnL", "numparams", free_params, "AICc", "AICc_wt")

# Also save to text files
write.table(restable_AIC_rellike, file="restablew_AIC_rellikec.txt", quote=FALSE, sep="\t")
write.table(restable_AICc_rellike, file="restablew_AICc_rellikec.txt", quote=FALSE, sep="\t")

# Save with nice conditional formatting
write.table(conditional_format_table(restable_AIC_rellike), file="restablew_AIC_rellike_formattedc.txt", quote=FALSE, sep="\t")
write.table(conditional_format_table(restable_AICc_rellike), file="restablew_AICc_rellike_formattedc.txt", quote=FALSE, sep="\t")


```


0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-
START HERE FOR FINAL ANALYSES
0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-

Now we will perform the same analyses with less biogeographical areas (Mediterranean - Med, Bubsaharian-Africa - AF, Assia-Far East including India, China, Japan... - AS, Australia and New Zealand - AU, North America - NAm, South America - SAm)

We think these further analyses are the most appropriate. Thus, we can discard previous analyses and keeping those below:


Redefine input occurrence data.

```{r, echo=F}

geogfn = np("Scleria_BGB_occurrence_world_bis.txt")
#extdata_dir = np(system.file("extdata", package="BioGeoBEARS"))
#extdata_dir
#list.files(extdata_dir)
#geogfn = np(paste(addslash(extdata_dir), "Psychotria_geog.data", sep=""))

geogfn
moref(geogfn)

# Look at your geographic range data:
tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
tipranges

max_range_size = 6

numstates_from_numareas(numareas=6, maxareas=6, include_null_range=TRUE)
#numstates_from_numareas(numareas=8, maxareas=8, include_null_range=FALSE)


```

17. Ahora vamos a usar el modelo DEC

```{r, echo=F}

#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Scleria_DEC_M0_unconstrained_v2.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDEC = res
    } else {
    # Loads to "res"
    load(resfn)
    resDEC = res
    }



```

18. Run DEC + J

```{r, echo=F}



#######################################################
# Run DEC+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDEC$outputs@params_table["d","est"]
estart = resDEC$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Add j as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
resfn = "Scleria_DEC+J_M0_unconstrained_v2.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDECj = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECj = res
    }
```


19. Plotting the results

```{r, echo=F}

#######################################################
# PDF plots
#######################################################
pdffn = "Scleria_DEC_vs_DEC+J_M0_unconstrained_v2.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC on Scleria M0_unconstrained_v2"

# Setup
results_object = resDEC
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges, plotlegend=TRUE)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges, plotlegend=TRUE)

#######################################################
# Plot ancestral states - DECJ
#######################################################
analysis_titletxt ="BioGeoBEARS DEC+J on Scleria M0_unconstrained_v2"

# Setup
results_object = resDECj
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges, plotlegend=TRUE)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges, plotlegend=TRUE)

dev.off()  # Turn off PDF



```

20. Run DIVA like

```{r, echo=F}
#######################################################
# Run DIVALIKE
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE model
# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# No jump dispersal/founder-event speciation
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.01
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.01

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

runslow = TRUE
resfn = "Scleria_DIVALIKE_M0_unconstrained_v2.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDIVALIKE = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKE = res
    }

```

21. Run DIVALIKE + j

```{r, echo=F}
#######################################################
# Run DIVALIKE+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDIVALIKE$outputs@params_table["d","est"]
estart = resDIVALIKE$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add jump dispersal/founder-event speciation
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Under DIVALIKE+J, the max of "j" should be 2, not 3 (as is default in DEC+J)
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","min"] = 0.00001
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","max"] = 1.99999

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

resfn = "Scleria_DIVALIKE+J_M0_unconstrained_v2.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDIVALIKEj = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEj = res
    }


```

22. Plotting additional results

```{r, echo=F}

pdffn = "Scleria_DIVALIKE_vs_DIVALIKE+J_M0_unconstrained_v2.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DIVALIKE
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE on Scleria M0_unconstrained_v2"

# Setup
results_object = resDIVALIKE
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DIVALIKE+J
#######################################################
analysis_titletxt ="Scleria DIVALIKE+J on Scleria M0_unconstrained_v2"

# Setup
results_object = resDIVALIKEj
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()



```



23. Comparing all results

```{r, echo=F}
#########################################################################
#########################################################################
#########################################################################
#########################################################################
# 
# CALCULATE SUMMARY STATISTICS TO COMPARE
# DEC, DEC+J, DIVALIKE, DIVALIKE+J
# 
#########################################################################
#########################################################################
#########################################################################
#########################################################################

#########################################################################
#########################################################################
# REQUIRED READING:
#
# Practical advice / notes / basic principles on statistical model 
#    comparison in general, and in BioGeoBEARS:
# http://phylo.wikidot.com/advice-on-statistical-model-comparison-in-biogeobears
#########################################################################
#########################################################################

# Set up empty tables to hold the statistical results
restable = NULL
teststable = NULL

#######################################################
# Statistics -- DEC vs. DEC+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDEC)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDECj)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DEC, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDEC, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DEC+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECj, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

# The null hypothesis for a Likelihood Ratio Test (LRT) is that two models
# confer the same likelihood on the data. See: Brian O'Meara's webpage:
# http://www.brianomeara.info/tutorials/aic
# ...for an intro to LRT, AIC, and AICc

rbind(res2, res1)
tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#######################################################
# Statistics -- DIVALIKE vs. DIVALIKE+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKE)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEj)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DIVALIKE, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKE, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DIVALIKE+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEj, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

rbind(res2, res1)
conditional_format_table(stats)

tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)


#########################################################################
# RESULTS: DEC, DEC+J, DIVALIKE, DIVALIKE+J
#########################################################################
teststable$alt = c("DEC+J", "DIVALIKE+J")
teststable$null = c("DEC", "DIVALIKE")
row.names(restable) = c("DEC", "DEC+J", "DIVALIKE", "DIVALIKE+J")

# Look at the results!!
restable
teststable

#######################################################
# Save the results tables for later -- check for e.g.
# convergence issues
#######################################################

# Loads to "restable"
save(restable, file="restable_v2.Rdata")
load(file="restable_v2.Rdata")

# Loads to "teststable"
save(teststable, file="teststable_v2.Rdata")
load(file="teststable_v2.Rdata")

# Also save to text files
write.table(restable, file="restable_v2.txt", quote=FALSE, sep="\t")
write.table(unlist_df(teststable), file="teststable_v2.txt", quote=FALSE, sep="\t")

#######################################################
# Model weights of all four models
#######################################################
restable2 = restable

# With AICs:
AICtable = calc_AIC_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams)
restable = cbind(restable, AICtable)
restable_AIC_rellike = AkaikeWeights_on_summary_table(restable=restable, colname_to_use="AIC")
restable_AIC_rellike

# With AICcs -- factors in sample size
samplesize = length(tr$tip.label)
AICtable = calc_AICc_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams, samplesize=samplesize)
restable2 = cbind(restable2, AICtable)
restable_AICc_rellike = AkaikeWeights_on_summary_table(restable=restable2, colname_to_use="AIC")
restable_AICc_rellike
free_params = row.names(resDECj$output@params_table[resDECj$output@params_table$type=="free",])
names(restable_AICc_rellike) = c("LnL", "numparams", free_params, "AICc", "AICc_wt")

# Also save to text files
write.table(restable_AIC_rellike, file="restable_v2_AIC_rellike.txt", quote=FALSE, sep="\t")
write.table(restable_AICc_rellike, file="restable_v2_AICc_rellike.txt", quote=FALSE, sep="\t")

# Save with nice conditional formatting
write.table(conditional_format_table(restable_AIC_rellike), file="restable_v2_AIC_rellike_formatted.txt", quote=FALSE, sep="\t")
write.table(conditional_format_table(restable_AICc_rellike), file="restable_v2_AICc_rellike_formatted.txt", quote=FALSE, sep="\t")


```


End of the first approach: DEC, DEC+J, DIVALIKE and DIVALIKE+J without dispersal multipliers.




24. Now we are going to use the model DEC with dispersal multipliers allowing the dispersal matrix to be modified by extra parameter -w

```{r, echo=F}

#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers_bis.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = areasadjacency
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Scleria_DEC_M0_constrained_w_v2.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDECc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECc = res
    }



```



25. Run DEC + J con tasas de dispersion

```{r, echo=F}



#######################################################
# Run DEC+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers_bis.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDEC$outputs@params_table["d","est"]
estart = resDEC$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Add j as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
resfn = "Scleria_DEC+J_M0_constrained_w_v2.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDECjc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECjc = res
    }
```


26. Plotting the results with dispesal rates

```{r, echo=F}

#######################################################
# PDF plots
#######################################################
pdffn = "Scleria_DEC_vs_DEC+J_M0_constrained_w_v2.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC on Scleria M0_constrained_v2"

# Setup
results_object = resDECc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DECJ
#######################################################
analysis_titletxt ="BioGeoBEARS DEC+J on Scleria M0_constrained_v2"

# Setup
results_object = resDECjc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()  # Turn off PDF

 # Plot it

```

27. Run DIVA like with dispersal rates

```{r, echo=F}
#######################################################
# Run DIVALIKE
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers_bis.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE model
# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

# No jump dispersal/founder-event speciation
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.01
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.01

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

runslow = TRUE
resfn = "Scleria_DIVALIKE_M0_constrained_w_v2.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDIVALIKEc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEc = res
    }

```

28. Run DIVALIKE + j with dispersal rates

```{r, echo=F}
#######################################################
# Run DIVALIKE+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers_bis.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDIVALIKE$outputs@params_table["d","est"]
estart = resDIVALIKE$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add jump dispersal/founder-event speciation
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Under DIVALIKE+J, the max of "j" should be 2, not 3 (as is default in DEC+J)
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","min"] = 0.00001
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","max"] = 1.99999

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

resfn = "Scleria_DIVALIKE+J_M0_constrained_w_v2.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDIVALIKEjc = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEjc = res
    }


```

29. Plotting additional results with dispersal rates

```{r, echo=F}

pdffn = "Scleria_DIVALIKE_vs_DIVALIKE+J_M0_constrained_w_v2.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DIVALIKE
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE on Scleria M0_constrained_v2"

# Setup
results_object = resDIVALIKEc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DIVALIKE+J
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE+J on Scleria M0_constrained_v2"

# Setup
results_object = resDIVALIKEjc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()



```



30. Comparing all results with dispersal rates

```{r, echo=F}
#########################################################################
#########################################################################
#########################################################################
#########################################################################
# 
# CALCULATE SUMMARY STATISTICS TO COMPARE
# DEC, DEC+J, DIVALIKE, DIVALIKE+J
# 
#########################################################################
#########################################################################
#########################################################################
#########################################################################

#########################################################################
#########################################################################
# REQUIRED READING:
#
# Practical advice / notes / basic principles on statistical model 
#    comparison in general, and in BioGeoBEARS:
# http://phylo.wikidot.com/advice-on-statistical-model-comparison-in-biogeobears
#########################################################################
#########################################################################

# Set up empty tables to hold the statistical results
restable = NULL
teststable = NULL

#######################################################
# Statistics -- DEC vs. DEC+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDECc)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDECjc)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DEC, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DEC+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECjc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

# The null hypothesis for a Likelihood Ratio Test (LRT) is that two models
# confer the same likelihood on the data. See: Brian O'Meara's webpage:
# http://www.brianomeara.info/tutorials/aic
# ...for an intro to LRT, AIC, and AICc

rbind(res2, res1)
tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#######################################################
# Statistics -- DIVALIKE vs. DIVALIKE+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEc)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEjc)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DIVALIKE, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DIVALIKE+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEjc, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

rbind(res2, res1)
conditional_format_table(stats)

tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#########################################################################
# RESULTS: DEC, DEC+J, DIVALIKE, DIVALIKE+J
#########################################################################
teststable$alt = c("DEC+Jc", "DIVALIKE+Jc")
teststable$null = c("DECc", "DIVALIKEc")
row.names(restable) = c("DECc", "DEC+Jc", "DIVALIKEc", "DIVALIKE+Jc")

# Look at the results!!
restable
teststable

#######################################################
# Save the results tables for later -- check for e.g.
# convergence issues
#######################################################

# Loads to "restable"
save(restable, file="restable_wc_v2.Rdata")
load(file="restable_wc_v2.Rdata")

# Loads to "teststable"
save(teststable, file="teststable_wc_v2.Rdata")
load(file="teststable_wc.Rdata")

# Also save to text files
write.table(restable, file="restablewc_v2.txt", quote=FALSE, sep="\t")
write.table(unlist_df(teststable), file="teststablewc_v2.txt", quote=FALSE, sep="\t")

#######################################################
# Model weights of all six models
#######################################################
restable2 = restable

# With AICs:
AICtable = calc_AIC_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams)
restable = cbind(restable, AICtable)
restable_AIC_rellike = AkaikeWeights_on_summary_table(restable=restable, colname_to_use="AIC")
restable_AIC_rellike

# With AICcs -- factors in sample size
samplesize = length(tr$tip.label)
AICtable = calc_AICc_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams, samplesize=samplesize)
restable2 = cbind(restable2, AICtable)
restable_AICc_rellike = AkaikeWeights_on_summary_table(restable=restable2, colname_to_use="AIC")
restable_AICc_rellike
free_params = row.names(resDECj$output@params_table[resDECj$output@params_table$type=="free",])
names(restable_AICc_rellike) = c("LnL", "numparams", free_params, "AICc", "AICc_wt")

# Also save to text files
write.table(restable_AIC_rellike, file="restablew_AIC_rellikec_v2.txt", quote=FALSE, sep="\t")
write.table(restable_AICc_rellike, file="restablew_AICc_rellikec_v2.txt", quote=FALSE, sep="\t")

# Save with nice conditional formatting
write.table(conditional_format_table(restable_AIC_rellike), file="restablew_v2_AIC_rellike_formattedc.txt", quote=FALSE, sep="\t")
write.table(conditional_format_table(restable_AICc_rellike), file="restablew_v2_AICc_rellike_formattedc.txt", quote=FALSE, sep="\t")


```


0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-

0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-

MEDITERRANEAN ANALYSES

These are the same analyses, but with Mediterranean region divided into 4 sub-regions.The occurrence matrix is the same than those used in V2 (with less areas), splitting the Mediterranean region.


First, load the new distribution matrix.

```{r, echo=F}


geogfn = np("Scleria_BGB_occurrence_mediterranean_bis.txt")
#extdata_dir = np(system.file("extdata", package="BioGeoBEARS"))
#extdata_dir
#list.files(extdata_dir)
#geogfn = np(paste(addslash(extdata_dir), "Psychotria_geog.data", sep=""))

geogfn
moref(geogfn)

# Look at your geographic range data:
tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
tipranges

max_range_size = 9

numstates_from_numareas(numareas=9, maxareas=9, include_null_range=TRUE)
#numstates_from_numareas(numareas=8, maxareas=8, include_null_range=FALSE)


```



31. Ahora vamos a usar el modelo DEC

```{r, echo=F}

#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Scleria_DEC_M0_unconstrained_mediterranean.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDECm = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECm = res
    }



```



32. Run DEC + J

```{r, echo=F}



#######################################################
# Run DEC+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDEC$outputs@params_table["d","est"]
estart = resDEC$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Add j as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
resfn = "Scleria_DEC+J_M0_unconstrained_mediterranean.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDECjm = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECjm = res
    }
```


33. Plotting the results

```{r, echo=F}

#######################################################
# PDF plots
#######################################################
pdffn = "Scleria_DEC_vs_DEC+J_M0_unconstrained_mediterranean.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC on Scleria M0_unconstrained_mediterranean"

# Setup
results_object = resDECm
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DECJ
#######################################################
analysis_titletxt ="BioGeoBEARS DEC+J on Scleria M0_unconstrained_mediterranean"

# Setup
results_object = resDECjm
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()  # Turn off PDF



```



34. Run DIVA like

```{r, echo=F}
#######################################################
# Run DIVALIKE
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE model
# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# No jump dispersal/founder-event speciation
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.01
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.01

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

runslow = TRUE
resfn = "Scleria_DIVALIKE_M0_unconstrained_mediterranean.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDIVALIKEm = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEm = res
    }

```

35. Run DIVALIKE + j

```{r, echo=F}
#######################################################
# Run DIVALIKE+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
#BioGeoBEARS_run_object$timesfn = "timeperiods.txt"
#BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
#BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDIVALIKE$outputs@params_table["d","est"]
estart = resDIVALIKE$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add jump dispersal/founder-event speciation
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Under DIVALIKE+J, the max of "j" should be 2, not 3 (as is default in DEC+J)
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","min"] = 0.00001
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","max"] = 1.99999

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

resfn = "Scleria_DIVALIKE+J_M0_unconstrained_mediterranean.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDIVALIKEjm = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEjm = res
    }


```

36. Plotting additional results

```{r, echo=F}

pdffn = "Scleria_DIVALIKE_vs_DIVALIKE+J_M0_unconstrained_mediterranean.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DIVALIKE
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE on Scleria M0_unconstrained_mediterranean"

# Setup
results_object = resDIVALIKEm
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DIVALIKE+J
#######################################################
analysis_titletxt ="Scleria DIVALIKE+J on Scleria M0_unconstrained_mediterranean"

# Setup
results_object = resDIVALIKEjm
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()



```



37. Comparing all results

```{r, echo=F}
#########################################################################
#########################################################################
#########################################################################
#########################################################################
# 
# CALCULATE SUMMARY STATISTICS TO COMPARE
# DEC, DEC+J, DIVALIKE, DIVALIKE+J
# 
#########################################################################
#########################################################################
#########################################################################
#########################################################################

#########################################################################
#########################################################################
# REQUIRED READING:
#
# Practical advice / notes / basic principles on statistical model 
#    comparison in general, and in BioGeoBEARS:
# http://phylo.wikidot.com/advice-on-statistical-model-comparison-in-biogeobears
#########################################################################
#########################################################################

# Set up empty tables to hold the statistical results
restable = NULL
teststable = NULL

#######################################################
# Statistics -- DEC vs. DEC+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDECm)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDECjm)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DEC, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECm, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DEC+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECjm, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

# The null hypothesis for a Likelihood Ratio Test (LRT) is that two models
# confer the same likelihood on the data. See: Brian O'Meara's webpage:
# http://www.brianomeara.info/tutorials/aic
# ...for an intro to LRT, AIC, and AICc

rbind(res2, res1)
tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#######################################################
# Statistics -- DIVALIKE vs. DIVALIKE+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEm)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEjm)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DIVALIKE, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEm, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DIVALIKE+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEjm, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

rbind(res2, res1)
conditional_format_table(stats)

tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)


#########################################################################
# RESULTS: DEC, DEC+J, DIVALIKE, DIVALIKE+J
#########################################################################
teststable$alt = c("DEC+J", "DIVALIKE+J")
teststable$null = c("DEC", "DIVALIKE")
row.names(restable) = c("DEC", "DEC+J", "DIVALIKE", "DIVALIKE+J")

# Look at the results!!
restable
teststable

#######################################################
# Save the results tables for later -- check for e.g.
# convergence issues
#######################################################

# Loads to "restable"
save(restable, file="restable_mediterranean.Rdata")
load(file="restable_mediterranean.Rdata")

# Loads to "teststable"
save(teststable, file="teststable_mediterranean.Rdata")
load(file="teststable_mediterranean.Rdata")

# Also save to text files
write.table(restable, file="restable_mediterranean.txt", quote=FALSE, sep="\t")
write.table(unlist_df(teststable), file="teststable_mediterranean.txt", quote=FALSE, sep="\t")

#######################################################
# Model weights of all four models
#######################################################
restable2 = restable

# With AICs:
AICtable = calc_AIC_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams)
restable = cbind(restable, AICtable)
restable_AIC_rellike = AkaikeWeights_on_summary_table(restable=restable, colname_to_use="AIC")
restable_AIC_rellike

# With AICcs -- factors in sample size
samplesize = length(tr$tip.label)
AICtable = calc_AICc_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams, samplesize=samplesize)
restable2 = cbind(restable2, AICtable)
restable_AICc_rellike = AkaikeWeights_on_summary_table(restable=restable2, colname_to_use="AIC")
restable_AICc_rellike
free_params = row.names(resDECjm$output@params_table[resDECjm$output@params_table$type=="free",])
names(restable_AICc_rellike) = c("LnL", "numparams", free_params, "AICc", "AICc_wt")

# Also save to text files
write.table(restable_AIC_rellike, file="restable_mediterranean_AIC_rellike.txt", quote=FALSE, sep="\t")
write.table(restable_AICc_rellike, file="restable_mediterranean_AICc_rellike.txt", quote=FALSE, sep="\t")

# Save with nice conditional formatting
write.table(conditional_format_table(restable_AIC_rellike), file="restable_mediterranean_AIC_rellike_formatted.txt", quote=FALSE, sep="\t")
write.table(conditional_format_table(restable_AICc_rellike), file="restable_mediterranean_AICc_rellike_formatted.txt", quote=FALSE, sep="\t")


```


End of the first approach: DEC, DEC+J, DIVALIKE and DIVALIKE+J without dispersal multipliers.


--------------07.02.2019----------


38. Now we are going to use the model DEC with dispersal multipliers allowing the dispersal matrix to be modified by extra parameter -w

```{r, echo=F}

#######################################################
# Run DEC
#######################################################

# Intitialize a default model (DEC model)
BioGeoBEARS_run_object = define_BioGeoBEARS_run()

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
# 1. Here, un-comment ONLY the files you want to use.
# 2. Also un-comment "BioGeoBEARS_run_object = section_the_tree(...", below.
# 3. For example files see (a) extdata_dir, 
#  or (b) http://phylo.wikidot.com/biogeobears#files
#  and BioGeoBEARS Google Group posts for further hints)
#
# Uncomment files you wish to use in time-stratified analyses:
BioGeoBEARS_run_object$timesfn = "timeperiods_med.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers_mediterranean.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = areasadjacency
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC model
# (nothing to do; defaults)

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Scleria_DEC_M0_constrained_w_mediterranean.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDECcm = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECcm = res
    }



```



39. Run DEC + J con tasas de dispersion

```{r, echo=F}



#######################################################
# Run DEC+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods_med.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers_mediterranean.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DEC+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDEC$outputs@params_table["d","est"]
estart = resDEC$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Add j as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
resfn = "Scleria_DEC+J_M0_constrained_w_mediterranean.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDECjcm = res
    } else {
    # Loads to "res"
    load(resfn)
    resDECjcm = res
    }
```



40. Plotting the results with dispesal rates

```{r, echo=F}

#######################################################
# PDF plots
#######################################################
pdffn = "Scleria_DEC_vs_DEC+J_M0_constrained_w_mediterranean.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DEC
#######################################################
analysis_titletxt ="BioGeoBEARS DEC on Scleria M0_constrained_med"

# Setup
results_object = resDECcm
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DECJ
#######################################################
analysis_titletxt ="BioGeoBEARS DEC+J on Scleria M0_constrained_med"

# Setup
results_object = resDECjcm
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()  # Turn off PDF

 # Plot it

```


27.01.2019-------------------

41. Run DIVA like with dispersal rates

```{r, echo=F}
#######################################################
# Run DIVALIKE
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods_med.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers_mediterranean.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE model
# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

# No jump dispersal/founder-event speciation
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.01
# BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.01

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

runslow = TRUE
resfn = "Scleria_DIVALIKE_M0_constrained_w_mediterranean.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    resDIVALIKEcm = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEcm = res
    }

```



42. Run DIVALIKE + j with dispersal rates

```{r, echo=F}
#######################################################
# Run DIVALIKE+J
#######################################################
BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$trfn = trfn
BioGeoBEARS_run_object$geogfn = geogfn
BioGeoBEARS_run_object$max_range_size = max_range_size
BioGeoBEARS_run_object$min_branchlength = 0.000001    # Min to treat tip as a direct ancestor (no speciation event)
BioGeoBEARS_run_object$include_null_range = TRUE    # set to FALSE for e.g. DEC* model, DEC*+J, etc. (see Massana et al.)
# also search script on "include_null_range" for other places to change

# Set up a time-stratified analysis:
BioGeoBEARS_run_object$timesfn = "timeperiods_med.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "Scleria_BGB_manual_dispersal_multipliers_mediterranean.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = "areas_allowed.txt"
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Speed options and multicore processing if desired
BioGeoBEARS_run_object$speedup = TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE     # if FALSE, use optim() instead of optimx()
BioGeoBEARS_run_object$num_cores_to_use = 1
BioGeoBEARS_run_object$force_sparse = FALSE    # force_sparse=TRUE causes pathology & isn't much faster at this scale

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE    # get ancestral states from optim run

# Set up DIVALIKE+J model
# Get the ML parameter values from the 2-parameter nested model
# (this will ensure that the 3-parameter model always does at least as good)
dstart = resDIVALIKE$outputs@params_table["d","est"]
estart = resDIVALIKE$outputs@params_table["e","est"]
jstart = 0.0001

# Input starting values for d, e
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","init"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["d","est"] = dstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","init"] = estart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["e","est"] = estart

# Remove subset-sympatry
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","init"] = 0.0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["s","est"] = 0.0

BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ysv","type"] = "2-j"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["ys","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["y","type"] = "ysv*1/2"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["v","type"] = "ysv*1/2"

# Allow classic, widespread vicariance; all events equiprobable
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","type"] = "fixed"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","init"] = 0.5
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["mx01v","est"] = 0.5

# Add jump dispersal/founder-event speciation
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = jstart
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = jstart

# Under DIVALIKE+J, the max of "j" should be 2, not 3 (as is default in DEC+J)
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","min"] = 0.00001
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","max"] = 1.99999

# Add w as a free parameter
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","type"] = "free"
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","init"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","est"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","min"] = 0
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["w","max"] = 3

check_BioGeoBEARS_run(BioGeoBEARS_run_object)

resfn = "Scleria_DIVALIKE+J_M0_constrained_w_mediterranean.Rdata"
runslow = TRUE
if (runslow)
    {
    #sourceall("/Dropbox/_njm/__packages/BioGeoBEARS_setup/")

    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)

    resDIVALIKEjcm = res
    } else {
    # Loads to "res"
    load(resfn)
    resDIVALIKEjcm = res
    }


```



43. Plotting additional results with dispersal rates

```{r, echo=F}

pdffn = "Scleria_DIVALIKE_vs_DIVALIKE+J_M0_constrained_w_mediterranean.pdf"
pdf(pdffn, width=7, height=9)

#######################################################
# Plot ancestral states - DIVALIKE
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE on Scleria M0_constrained_med"

# Setup
results_object = resDIVALIKEcm
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

#######################################################
# Plot ancestral states - DIVALIKE+J
#######################################################
analysis_titletxt ="BioGeoBEARS DIVALIKE+J on Scleria M0_constrained_med"

# Setup
results_object = resDIVALIKEjc
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res1 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.5, tipcex=0.4, statecex=0.2, splitcex=0.2, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.5, tipcex=0.4, statecex=0.5, splitcex=0.5, titlecex=0.9, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()



```



44. Comparing all results with dispersal rates

```{r, echo=F}
#########################################################################
#########################################################################
#########################################################################
#########################################################################
# 
# CALCULATE SUMMARY STATISTICS TO COMPARE
# DEC, DEC+J, DIVALIKE, DIVALIKE+J
# 
#########################################################################
#########################################################################
#########################################################################
#########################################################################

#########################################################################
#########################################################################
# REQUIRED READING:
#
# Practical advice / notes / basic principles on statistical model 
#    comparison in general, and in BioGeoBEARS:
# http://phylo.wikidot.com/advice-on-statistical-model-comparison-in-biogeobears
#########################################################################
#########################################################################

# Set up empty tables to hold the statistical results
restable = NULL
teststable = NULL

#######################################################
# Statistics -- DEC vs. DEC+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDECcm)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDECjcm)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DEC, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECcm, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DEC+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDECjcm, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

# The null hypothesis for a Likelihood Ratio Test (LRT) is that two models
# confer the same likelihood on the data. See: Brian O'Meara's webpage:
# http://www.brianomeara.info/tutorials/aic
# ...for an intro to LRT, AIC, and AICc

rbind(res2, res1)
tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#######################################################
# Statistics -- DIVALIKE vs. DIVALIKE+J
#######################################################
# We have to extract the log-likelihood differently, depending on the 
# version of optim/optimx
LnL_2 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEcm)
LnL_1 = get_LnL_from_BioGeoBEARS_results_object(resDIVALIKEjcm)

numparams1 = 3
numparams2 = 2
stats = AICstats_2models(LnL_1, LnL_2, numparams1, numparams2)
stats

# DIVALIKE, null model for Likelihood Ratio Test (LRT)
res2 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEcm, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)
# DIVALIKE+J, alternative model for Likelihood Ratio Test (LRT)
res1 = extract_params_from_BioGeoBEARS_results_object(results_object=resDIVALIKEjcm, returnwhat="table", addl_params=c("j"), paramsstr_digits=4)

rbind(res2, res1)
conditional_format_table(stats)

tmp_tests = conditional_format_table(stats)

restable = rbind(restable, res2, res1)
teststable = rbind(teststable, tmp_tests)

#########################################################################
# RESULTS: DEC, DEC+J, DIVALIKE, DIVALIKE+J
#########################################################################
teststable$alt = c("DEC+Jc", "DIVALIKE+Jc")
teststable$null = c("DECc", "DIVALIKEc")
row.names(restable) = c("DECc", "DEC+Jc", "DIVALIKEc", "DIVALIKE+Jc")

# Look at the results!!
restable
teststable

#######################################################
# Save the results tables for later -- check for e.g.
# convergence issues
#######################################################

# Loads to "restable"
save(restable, file="restable_wc_med.Rdata")
load(file="restable_wc_med.Rdata")

# Loads to "teststable"
save(teststable, file="teststable_wc_med.Rdata")
load(file="teststable_wc_med.Rdata")

# Also save to text files
write.table(restable, file="restablewc_med.txt", quote=FALSE, sep="\t")
write.table(unlist_df(teststable), file="teststablewc_med.txt", quote=FALSE, sep="\t")

#######################################################
# Model weights of all six models
#######################################################
restable2 = restable

# With AICs:
AICtable = calc_AIC_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams)
restable = cbind(restable, AICtable)
restable_AIC_rellike = AkaikeWeights_on_summary_table(restable=restable, colname_to_use="AIC")
restable_AIC_rellike

# With AICcs -- factors in sample size
samplesize = length(tr$tip.label)
AICtable = calc_AICc_column(LnL_vals=restable$LnL, nparam_vals=restable$numparams, samplesize=samplesize)
restable2 = cbind(restable2, AICtable)
restable_AICc_rellike = AkaikeWeights_on_summary_table(restable=restable2, colname_to_use="AIC")
restable_AICc_rellike
free_params = row.names(resDECjm$output@params_table[resDECjm$output@params_table$type=="free",])
names(restable_AICc_rellike) = c("LnL", "numparams", free_params, "AICc", "AICc_wt")

# Also save to text files
write.table(restable_AIC_rellike, file="restablew_AIC_rellikec_med.txt", quote=FALSE, sep="\t")
write.table(restable_AICc_rellike, file="restablew_AICc_rellikec_med.txt", quote=FALSE, sep="\t")

# Save with nice conditional formatting
write.table(conditional_format_table(restable_AIC_rellike), file="restablew_AIC_rellike_formattedc_med.txt", quote=FALSE, sep="\t")
write.table(conditional_format_table(restable_AICc_rellike), file="restablew_AICc_rellike_formattedc_med.txt", quote=FALSE, sep="\t")


```


0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-
31.01.2019
0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-0-


End of first approach.




-----11.02.2019--------




35. BAYESIAN STOCHASTIC MAPPING

```{r, echo=F}


# Load the package (after installation, see above).
library(optimx)         # You need to have some version of optimx available
                        # as it is a BioGeoBEARS dependency; however, if you
                        # don't want to use optimx, and use optim() (from R core) 
                        # you can set:
                        # BioGeoBEARS_run_object$use_optimx = FALSE
                        # ...everything should work either way -- NJM 2014-01-08
library(FD)       # for FD::maxent() (make sure this is up-to-date)
library(snow)     # (if you want to use multicore functionality; some systems/R versions prefer library(parallel), try either)
library(parallel)
library(BioGeoBEARS)

########################################################
# TO GET THE OPTIMX/OPTIM FIX, AND THE UPPASS FIX, 
# SOURCE THE REVISED FUNCTIONS WITH THESE COMMANDS
#
# CRUCIAL CRUCIAL CRUCIAL: 
# YOU HAVE TO RUN THE SOURCE COMMANDS AFTER 
# *EVERY TIME* YOU DO library(BioGeoBEARS). THE CHANGES ARE NOT "PERMANENT", 
# THEY HAVE TO BE MADE EACH TIME.  IF YOU ARE GOING TO BE OFFLINE, 
# YOU CAN DOWNLOAD EACH .R FILE TO YOUR HARD DRIVE AND REFER THE source()
# COMMANDS TO THE FULL PATH AND FILENAME OF EACH FILE ON YOUR
# LOCAL SYSTEM INSTEAD.
########################################################
library(BioGeoBEARS)
library(roxygen2)
source("http://phylo.wdfiles.com/local--files/biogeobears/cladoRcpp.R") # (needed now that traits model added; source FIRST!)
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_add_fossils_randomly_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_basics_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_calc_transition_matrices_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_classes_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_detection_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_DNA_cladogenesis_sim_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_extract_Qmat_COOmat_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_generics_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_models_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_on_multiple_trees_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_plots_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_readwrite_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_simulate_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_SSEsim_makePlots_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_SSEsim_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_stochastic_mapping_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_stratified_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_univ_model_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/calc_uppass_probs_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/calc_loglike_sp_v01.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/get_stratified_subbranch_top_downpass_likelihoods_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/runBSM_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/stochastic_map_given_inputs.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/summarize_BSM_tables_v1.R")
source("http://phylo.wdfiles.com/local--files/biogeobears/BioGeoBEARS_traits_v1.R") # added traits model
calc_loglike_sp = compiler::cmpfun(calc_loglike_sp_prebyte)    # crucial to fix bug in uppass calculations
calc_independent_likelihoods_on_each_branch = compiler::cmpfun(calc_independent_likelihoods_on_each_branch_prebyte)
    # slight speedup hopefully

#######################################################
# Local source()-ing method -- uses BioGeoBEARS sourceall() function 
# on a directory of .R files, so you don't have to type them out.
# The directories here are on my machine, you would have to make a 
# directory, save the .R files there, and refer to them.
#
# NOTE: it's best to source the "cladoRcpp.R" update first, to avoid warnings like this:
##
## Note: possible error in 'rcpp_calc_anclikes_sp_COOweights_faster(Rcpp_leftprobs = tmpca_1, ': 
##         unused arguments (m = m, m_null_range = include_null_range, jts_matrix = jts_matrix) 
##
#
# TO USE: Delete or comment out the 'source("http://...")' commands above, and un-comment
#              the below...
########################################################################
# Un-comment (and fix directory paths) to use:
#library(BioGeoBEARS)
#source("/drives/Dropbox/_njm/__packages/cladoRcpp_setup/cladoRcpp.R")
#sourceall("/drives/Dropbox/_njm/__packages/BioGeoBEARS_setup/")
#calc_loglike_sp = compiler::cmpfun(calc_loglike_sp_prebyte)    # crucial to fix bug in uppass calculations
#calc_independent_likelihoods_on_each_branch = compiler::cmpfun(calc_independent_likelihoods_on_each_branch_prebyte)
########################################################################

#######################################################
# SETUP: YOUR WORKING DIRECTORY
#######################################################
# You will need to set your working directory to match your local system

# Note these very handy functions!
# Command "setwd(x)" sets your working directory
# Command "getwd()" gets your working directory and tells you what it is.
# Command "list.files()" lists the files in your working directory
# To get help on any command, use "?".  E.g., "?list.files"

# Set your working directory for output files
# default here is your home directory ("~")
# Change this as you like
#wd = "/drives/Dropbox/_njm/__packages/BioGeoBEARS_setup/inst/extdata/examples/Psychotria_M0/BGB_BSM_DEC+J_M3b_strat/"
#setwd(wd)

# Double-check your working directory with getwd()
#getwd()

#######################################################
# SETUP: Extension data directory
#######################################################
# When R packages contain extra files, they are stored in the "extdata" directory 
# inside the installed package.
#
# BioGeoBEARS contains various example files and scripts in its extdata directory.
# 
# Each computer operating system might install BioGeoBEARS in a different place, 
# depending on your OS and settings. 
# 
# However, you can find the extdata directory like this:
##extdata_dir = np(system.file("extdata", package="BioGeoBEARS"))
##extdata_dir
##list.files(extdata_dir)

# Time-stratified files are here
##extdata_dir2 = np(slashslash(paste(extdata_dir, "examples/Psychotria_M3strat/BGB/", sep="/")))

# "system.file" looks in the directory of a specified package (in this case BioGeoBEARS)
# The function "np" is just a shortcut for normalizePath(), which converts the 
# path to the format appropriate for your system (e.g., Mac/Linux use "/", but 
# Windows uses "\\", if memory serves).

# Even when using your own data files, you should KEEP these commands in your 
# script, since the plot_BioGeoBEARS_results function needs a script from the 
# extdata directory to calculate the positions of "corners" on the plot. This cannot
# be made into a straight up BioGeoBEARS function because it uses C routines 
# from the package APE which do not pass R CMD check for some reason.

#######################################################
# SETUP: YOUR TREE FILE AND GEOGRAPHY FILE
#######################################################
# Example files are given below. To run your own data,
# make the below lines point to your own files, e.g.
# trfn = "/mydata/frogs/frogBGB/tree.newick"
# geogfn = "/mydata/frogs/frogBGB/geog.data"

#######################################################
# Phylogeny file
# Notes: 
# 1. Must be binary/bifurcating: no polytomies
# 2. No negative branchlengths (e.g. BEAST MCC consensus trees sometimes have negative branchlengths)
# 3. Be careful of very short branches, as BioGeoBEARS will interpret ultrashort branches as direct ancestors
# 4. You can use non-ultrametric trees, but BioGeoBEARS will interpret any tips significantly below the 
#    top of the tree as fossils!  This is only a good idea if you actually do have fossils in your tree,
#    as in e.g. Wood, Matzke et al. (2013), Systematic Biology.
# 5. The default settings of BioGeoBEARS make sense for trees where the branchlengths are in units of 
#    millions of years, and the tree is 1-1000 units tall. If you have a tree with a total height of
#    e.g. 0.00001, you will need to adjust e.g. the max values of d and e, or (simpler) multiply all
#    your branchlengths to get them into reasonable units.
# 6. DON'T USE SPACES IN SPECIES NAMES, USE E.G. "_"
#######################################################
# This is the example Newick file for Hawaiian JuliaDupin
# (from Ree & Smith 2008)
# "trfn" = "tree file name"
trfn = np("glareosae.tree")

# Look at the raw Newick file:
moref(trfn)

# Look at your phylogeny:
tr = read.tree(trfn)
tr
#plot(tr)
#title("Florestina phylogeny")
#axisPhylo() # plots timescale

#######################################################
# Geography file
# Notes:
# 1. This is a PHLYIP-formatted file. This means that in the 
#    first line, 
#    - the 1st number equals the number of rows (species)
#    - the 2nd number equals the number of columns (number of areas)
# 2. This is the same format used for C++ LAGRANGE geography files.
# 3. All names in the geography file must match names in the phylogeny file.
# 4. DON'T USE SPACES IN SPECIES NAMES, USE E.G. "_"
# 5. Operational taxonomic units (OTUs) should ideally be phylogenetic lineages, 
#    i.e. genetically isolated populations.  These may or may not be identical 
#    with species.  You would NOT want to just use specimens, as each specimen 
#    automatically can only live in 1 area, which will typically favor DEC+J 
#    models.  This is fine if the species/lineages really do live in single areas,
#    but you wouldn't want to assume this without thinking about it at least. 
#    In summary, you should collapse multiple specimens into species/lineages if 
#    data indicates they are the same genetic population.
######################################################

# This is the example geography file for Hawaiian JuliaDupin
# (from Ree & Smith 2008)
geogfn = np(paste("glareosae2.data"))

# Look at the raw geography text file:
moref(geogfn)

# Look at your geographic range data:
tipranges = getranges_from_LagrangePHYLIP(lgdata_fn=geogfn)
tipranges

# Set the maximum number of areas any species may occupy; this cannot be larger 
# than the number of areas you set up, but it can be smaller.
max_range_size = 5

#######################################################
# Run DEC
#######################################################

BioGeoBEARS_run_object = define_BioGeoBEARS_run()
BioGeoBEARS_run_object$force_sparse=FALSE    # sparse=FALSE causes pathology & isn't much faster at this scale
BioGeoBEARS_run_object$speedup=TRUE          # shorcuts to speed ML search; use FALSE if worried (e.g. >3 params)
BioGeoBEARS_run_object$use_optimx = TRUE
BioGeoBEARS_run_object$calc_ancprobs=TRUE    # get ancestral states from optim run

# Set up a time-stratified analysis 
# (un-comment to use; see example files in extdata_dir, 
#  and BioGeoBEARS google group posts for further hints)
BioGeoBEARS_run_object$timesfn = "timeperiods2.txt"
BioGeoBEARS_run_object$dispersal_multipliers_fn = "manual_dispersal_multipliers2b.txt"
#BioGeoBEARS_run_object$areas_allowed_fn = np(slashslash(paste(extdata_dir2, "areas_allowed.txt", sep="/")))
#BioGeoBEARS_run_object$areas_adjacency_fn = "areas_adjacency.txt"
#BioGeoBEARS_run_object$distsfn = "distances_matrix.txt"
# See notes on the distances model on PhyloWiki's BioGeoBEARS updates page.

# Input the maximum range size
BioGeoBEARS_run_object$max_range_size = max_range_size

# Multicore processing if desired
BioGeoBEARS_run_object$num_cores_to_use=1
# (use more cores to speed it up; this requires
# library(parallel) and/or library(snow). The package "parallel" 
# is now default on Macs in R 3.0+, but apparently still 
# has to be typed on some Windows machines. Note: apparently 
# parallel works on Mac command-line R, but not R.app.
# BioGeoBEARS checks for this and resets to 1
# core with R.app)

# Sparse matrix exponentiation is an option for huge numbers of ranges/states (600+)
# I have experimented with sparse matrix exponentiation in EXPOKIT/rexpokit,
# but the results are imprecise and so I haven't explored it further.
# In a Bayesian analysis, it might work OK, but the ML point estimates are
# not identical.
# Also, I have not implemented all functions to work with force_sparse=TRUE.
# Volunteers are welcome to work on it!!
BioGeoBEARS_run_object$force_sparse=FALSE

# Give BioGeoBEARS the location of the geography text file
BioGeoBEARS_run_object$geogfn = geogfn

# Give BioGeoBEARS the location of the phylogeny Newick file
BioGeoBEARS_run_object$trfn = trfn

# This function loads the dispersal multiplier matrix etc. from the text files into the model object. Required for these to work!
# (It also runs some checks on these inputs for certain errors.)
BioGeoBEARS_run_object = readfiles_BioGeoBEARS_run(BioGeoBEARS_run_object)

# Divide the tree up by timeperiods/strata (uncomment this for stratified analysis)
BioGeoBEARS_run_object = section_the_tree(inputs=BioGeoBEARS_run_object, make_master_table=TRUE, plot_pieces=FALSE)
# The stratified tree is described in this table:
#BioGeoBEARS_run_object$master_table

# Good default settings to get ancestral states
BioGeoBEARS_run_object$return_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_TTL_loglike_from_condlikes_table = TRUE
BioGeoBEARS_run_object$calc_ancprobs = TRUE

# Set up DEC+J model

# Add j as a free parameter
#BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","type"] = "free"
#BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","init"] = 0.001
#BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table["j","est"] = 0.001

# Look at the BioGeoBEARS_run_object; it's just a list of settings etc.
BioGeoBEARS_run_object

# This contains the model object
BioGeoBEARS_run_object$BioGeoBEARS_model_object

# This table contains the parameters of the model 
BioGeoBEARS_run_object$BioGeoBEARS_model_object@params_table

# Run this to check inputs. Read the error messages if you get them!
check_BioGeoBEARS_run(BioGeoBEARS_run_object)

# For a slow analysis, run once, then set runslow=FALSE to just 
# load the saved result.
runslow = TRUE
resfn = "Glareosae_DEC_ManDisMul_changethoughtime.Rdata"
if (runslow)
    {
    res = bears_optim_run(BioGeoBEARS_run_object)
    res    

    save(res, file=resfn)
    } else {
    # Loads to "res"
    load(resfn)
    }

###########################################
# Pick your model name:
###########################################
model_name = "DEC_ManDisMul_changethoughtime"
# res = resDEC

#######################################################
# Plot ancestral states - DEC
#######################################################
pdffn = paste0("Glareosae_", model_name, "_v1.pdf")
pdf(pdffn, width=6, height=6)

analysis_titletxt = paste0(model_name, " on Glareosae")

# Setup
results_object = res
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))

# States
res2 = plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="text", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

# Pie chart
plot_BioGeoBEARS_results(results_object, analysis_titletxt, addl_params=list("j"), plotwhat="pie", label.offset=0.45, tipcex=0.7, statecex=0.7, splitcex=0.6, titlecex=0.8, plotsplits=TRUE, cornercoords_loc=scriptdir, include_null_range=TRUE, tr=tr, tipranges=tipranges)

dev.off()  # Turn off PDF

 # Plot it

#######################################################
# Stochastic mapping on DEC M3b stratified with islands coming up
#######################################################
clado_events_tables = NULL
ana_events_tables = NULL
lnum = 0

#######################################################
# Get the inputs for Biogeographical Stochastic Mapping
# Note: this can be slow for large state spaces and trees, since 
# the independent likelihoods for each branch are being pre-calculated
# E.g., for 10 areas, this requires calculation of a 1024x1024 matrix
# for each branch.  On a tree with ~800 tips and thus ~1600 branches, this was about 1.6 gigs
# for storage of "BSM_inputs_file.Rdata".
# Update: 2015-09-23 -- now, if you used multicore functionality for the ML analysis,
# the same settings will be used for get_inputs_for_stochastic_mapping().
#######################################################
BSM_inputs_fn = "BSM_inputs_file.Rdata"
runInputsSlow = TRUE
if (runInputsSlow)
    {
    stochastic_mapping_inputs_list = get_inputs_for_stochastic_mapping(res=res)
    save(stochastic_mapping_inputs_list, file=BSM_inputs_fn)
    } else {
    # Loads to "stochastic_mapping_inputs_list"
    load(BSM_inputs_fn)
    } # END if (runInputsSlow)

# Check inputs (doesn't work the same on unconstr)
names(stochastic_mapping_inputs_list)
stochastic_mapping_inputs_list$phy2
stochastic_mapping_inputs_list$COO_weights_columnar
stochastic_mapping_inputs_list$unconstr
set.seed(seed=as.numeric(Sys.time()))

runBSMslow = TRUE
if (runBSMslow == TRUE)
    {
    # Saves to: RES_clado_events_tables.Rdata
    # Saves to: RES_ana_events_tables.Rdata
    BSM_output = runBSM(res, stochastic_mapping_inputs_list=stochastic_mapping_inputs_list, maxnum_maps_to_try=100, nummaps_goal=50, maxtries_per_branch=40000, save_after_every_try=TRUE, savedir=getwd(), seedval=12345, wait_before_save=0.01)

    RES_clado_events_tables = BSM_output$RES_clado_events_tables
    RES_ana_events_tables = BSM_output$RES_ana_events_tables
    } else {
    # Load previously saved...

    # Loads to: RES_clado_events_tables
    load(file="RES_clado_events_tables.Rdata")
    # Loads to: RES_ana_events_tables
    load(file="RES_ana_events_tables.Rdata")
    BSM_output = NULL
    BSM_output$RES_clado_events_tables = RES_clado_events_tables
    BSM_output$RES_ana_events_tables = RES_ana_events_tables
    } # END if (runBSMslow == TRUE)

# Extract BSM output
clado_events_tables = BSM_output$RES_clado_events_tables
ana_events_tables = BSM_output$RES_ana_events_tables
head(clado_events_tables[[1]])
head(ana_events_tables[[1]])
length(clado_events_tables)
length(ana_events_tables)

#######################################################
# Plot one stochastic map, manual method
#######################################################
# (we have to convert the stochastic maps into event
#  maps for plotting)

######################
# Get the color scheme
######################
include_null_range = TRUE
areanames = names(tipranges@df)
areas = areanames
max_range_size = 5

# Note: If you did something to change the states_list from the default given the number of areas, you would
# have to manually make that change here as well! (e.g., areas_allowed matrix, or manual reduction of the states_list)
states_list_0based = rcpp_areas_list_to_states_list(areas=areas, maxareas=max_range_size, include_null_range=include_null_range)

colors_list_for_states = get_colors_for_states_list_0based(areanames=areanames, states_list_0based=states_list_0based, max_range_size=max_range_size, plot_null_range=TRUE)

############################################
# Setup for painting a single stochastic map
############################################
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))
stratified=TRUE
clado_events_table = clado_events_tables[[1]]
ana_events_table = ana_events_tables[[1]]

# cols_to_get = names(clado_events_table[,-ncol(clado_events_table)])
# colnums = match(cols_to_get, names(ana_events_table))
# ana_events_table_cols_to_add = ana_events_table[,colnums]
# anagenetic_events_txt_below_node = rep("none", nrow(ana_events_table_cols_to_add))
# ana_events_table_cols_to_add = cbind(ana_events_table_cols_to_add, anagenetic_events_txt_below_node)
# rows_to_get_TF = ana_events_table_cols_to_add$node <= length(tr$tip.label)
# master_table_cladogenetic_events = rbind(ana_events_table_cols_to_add[rows_to_get_TF,], clado_events_table)

############################################
# Open a PDF
############################################
pdffn = paste0(model_name, "_single_stochastic_map_n1.pdf")
pdf(file=pdffn, width=6, height=6)

# Convert the BSM into a modified res object
master_table_cladogenetic_events = clado_events_tables[[1]]
resmod = stochastic_map_states_into_res(res=res, master_table_cladogenetic_events=master_table_cladogenetic_events, stratified=stratified)

plot_BioGeoBEARS_results(results_object=resmod, analysis_titletxt="Stochastic map", addl_params=list("j"), label.offset=0.5, plotwhat="text", cornercoords_loc=scriptdir, root.edge=TRUE, colors_list_for_states=colors_list_for_states, skiptree=FALSE, show.tip.label=TRUE)

# Paint on the branch states
paint_stochastic_map_branches(res=resmod, master_table_cladogenetic_events=master_table_cladogenetic_events, colors_list_for_states=colors_list_for_states, lwd=5, lty=par("lty"), root.edge=TRUE, stratified=stratified)

plot_BioGeoBEARS_results(results_object=resmod, analysis_titletxt="Stochastic map", addl_params=list("j"), plotwhat="text", cornercoords_loc=scriptdir, root.edge=TRUE, colors_list_for_states=colors_list_for_states, skiptree=TRUE, show.tip.label=TRUE)

############################################
# Close PDF
############################################
dev.off()
cmdstr = paste("xdg-open ", pdffn, sep="")


#######################################################
# Plot all 50 stochastic maps to PDF
#######################################################
# Setup
include_null_range = include_null_range
areanames = areanames
areas = areanames
max_range_size = max_range_size
states_list_0based = rcpp_areas_list_to_states_list(areas=areas, maxareas=max_range_size, include_null_range=include_null_range)
colors_list_for_states = get_colors_for_states_list_0based(areanames=areanames, states_list_0based=states_list_0based, max_range_size=max_range_size, plot_null_range=TRUE)
scriptdir = np(system.file("extdata/a_scripts", package="BioGeoBEARS"))
stratified = stratified

# Loop through the maps and plot to PDF
pdffn = paste0(model_name, "_", length(clado_events_tables), "BSMs_v1.pdf")
pdf(file=pdffn, width=6, height=6)

nummaps_goal = 50
for (i in 1:nummaps_goal)
    {
    clado_events_table = clado_events_tables[[i]]
    analysis_titletxt = paste0(model_name, " - Stochastic Map #", i, "/", nummaps_goal)
    plot_BSM(results_object=res, clado_events_table=clado_events_table, stratified=stratified, analysis_titletxt=analysis_titletxt, addl_params=list("j"), label.offset=0.5, plotwhat="text", cornercoords_loc=scriptdir, root.edge=TRUE, colors_list_for_states=colors_list_for_states, show.tip.label=TRUE, include_null_range=include_null_range)
    } # END for (i in 1:nummaps_goal)

dev.off()



#######################################################
# Summarize stochastic map tables
#######################################################
length(clado_events_tables)
length(ana_events_tables)

head(clado_events_tables[[1]][,-20])
tail(clado_events_tables[[1]][,-20])

head(ana_events_tables[[1]])
tail(ana_events_tables[[1]])

areanames = names(tipranges@df)
actual_names = areanames
actual_names

# Get the dmat and times (if any)
dmat_times = get_dmat_times_from_res(res=res, numstates=NULL)
dmat_times

# Extract BSM output
clado_events_tables = BSM_output$RES_clado_events_tables
ana_events_tables = BSM_output$RES_ana_events_tables

# Simulate the source areas
BSMs_w_sourceAreas = simulate_source_areas_ana_clado(res, clado_events_tables, ana_events_tables, areanames)
clado_events_tables = BSMs_w_sourceAreas$clado_events_tables
ana_events_tables = BSMs_w_sourceAreas$ana_events_tables

# Count all anagenetic and cladogenetic events
counts_list = count_ana_clado_events(clado_events_tables, ana_events_tables, areanames, actual_names)

summary_counts_BSMs = counts_list$summary_counts_BSMs
print(conditional_format_table(summary_counts_BSMs))

# Histogram of event counts
hist_event_counts(counts_list, pdffn=paste0(model_name, "_histograms_of_event_counts.pdf"))

#######################################################
# Print counts to files
#######################################################
tmpnames = names(counts_list)
cat("\n\nWriting tables* of counts to tab-delimited text files:\n(* = Tables have dimension=2 (rows and columns). Cubes (dimension 3) and lists (dimension 1) will not be printed to text files.) \n\n")
for (i in 1:length(tmpnames))
    {
    cmdtxt = paste0("item = counts_list$", tmpnames[i])
    eval(parse(text=cmdtxt))

    # Skip cubes
    if (length(dim(item)) != 2)
        {
        next()
        }

    outfn = paste0(tmpnames[i], ".txt")
    if (length(item) == 0)
        {
        cat(outfn, " -- NOT written, *NO* events recorded of this type", sep="")
        cat("\n")
        } else {
        cat(outfn)
        cat("\n")
        write.table(conditional_format_table(item), file=outfn, quote=FALSE, sep="\t", col.names=TRUE, row.names=TRUE)
        } # END if (length(item) == 0)
    } # END for (i in 1:length(tmpnames))
cat("...done.\n")

#######################################################
# Check that ML ancestral state/range probabilities and
# the mean of the BSMs approximately line up
#######################################################
library(MultinomialCI)    # For 95% CIs on BSM counts
check_ML_vs_BSM(res, clado_events_tables, model_name, tr=NULL, plot_each_node=FALSE, linreg_plot=TRUE, MultinomialCI=TRUE)
```

